%% This file is a part of fonts-tlwg package.

Fonts-TLWG is a collection of Thai scalable fonts available under free
licenses.  Its goal is to provide fonts that conform to existing standards
and recommendations, so that it can be a reference implementation.

It has been made available for use with pdfLaTeX.  To use it, you can use
the fonts-tlwg package by declaring like this in document preamble:

  \usepackage{fonts-tlwg}

OPTIONS
-------

Options are also available so that you can choose default fonts for use
in the document when switching to Thai in Babel:

* Using sans-serif font instead of roman:

  - sans       : Use sans-serif font by default when switching to Thai.

* Defining default roman, sans-serif, and teletype fonts:

  - rmkinnari  : Use "kinnari" as the default roman family.
  - rmnorasi   : Use "norasi" as the default roman family.
  - sfgaruda   : Use "garuda" as the default sans-serif family.
  - sflaksaman : Use "laksaman" as the default sans-serif family.
  - sfumpush   : Use "umpush" as the default sans-serif family.
  - sfloma     : Use "loma" as the default sans-serif family.
  - sfwaree    : Use "waree" as the default sans-serif family.
  - ttttype    : Use "ttype" as the default teletype family.
  - ttttypist  : Use "ttypist" as the default teletype family.

  These options do not change the default Thai font for the document.
  They only define the default Thai fonts of the three families for use
  in the document.

* Definging the default font for the document:

  - kinnari    : Use "kinnari" as the default font of the document.
  - garuda     : Use "garuda" as the default font of the document.
  - norasi     : Use "norasi" as the default font of the document.
  - laksaman   : Use "laksaman" as the default font of the document.
  - loma       : Use "loma" as the default font of the document.
  - purisa     : Use "purisa" as the default font of the document.
  - sawasdee   : Use "sawasdee" as the default font of the document.
  - ttype      : Use "ttype" as the default font of the document.
  - ttypist    : Use "ttypist" as the default font of the document.
  - umpush     : Use "umpush" as the default font of the document.
  - waree      : Use "waree" as the default font of the document.

  These options define the default Thai font when switching to Thai language
  in Babel.  They do not define fonts for the three families.


There is also another option for this package which would scale all Thai fonts
up or down by a given factor value. This is useful when there are discrepancies
between the size of the font for Latin script and the size of Thai fonts in
this package.

* Specifying the font scaling factor for every font in this package:

    - scale=<value>  : Set the scaling factor to the specified value.
                       The default value is 1.


Example:

* To use "laksaman" font all over the document (e.g. in Thai official letters):

    \usepackage[laksaman]{fonts-tlwg}

* To use "kinnari" as roman font instead of the default "norasi":

    \usepackage[rmkinnari]{fonts-tlwg}

* To typeset document with sans-serif font instead of roman:

    \usepackage[sans]{fonts-tlwg}

* To use "laksaman" as default font, allowing "kinnari" in roman text:

    \usepackage[sans,sflaksaman,rmkinnari]{fonts-tlwg}

  or:

    \usepackage[laksaman,rmkinnari]{fonts-tlwg}

* To use "laksaman" font with the size 20% larger than normal:

    \usepackage[laksaman,scaled=1.2]{fonts-tlwg}


\USEFONT COMMAND
----------------

You can still use the \usefont command where you want specific fonts in your
text portions, for example:

* To use "garuda" font with medium weight and normal shape:

    \usefont{LTH}{garuda}{m}{n}

* To use "kinnari" font with bold weight and italic shape:

    \usefont{LTH}{kinnari}{b}{it}

* To use "umpush" font with light weight (available for "umpush" only) and
  slant shape:

    \usefont{LTH}{umpush}{l}{sl}
